Pre Processing of the Short Term Data

### 2018-05-26 scripts/1_str_preprocessing.Rmd
Update to intro:
Update: Str Preprocessing Script to include relative paths.

---
title: "Pre Processing of the Short Term Data"
author: "Andy Krause"
date: "April 14, 2018"
output: html_document
---

This script executes a number of very long running data preparation process necessary to analyze the short term rental (Airbnb) data.  If you re-knit (with RMarkdown) this file the code will not evaluate.  To actually evaluate the code you must manually re-run the code chunks

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

@@

```{r load_libraries}

  library(plyr)
  library(stringr)
  library(lubridate)
  library(dplyr)
  library(multidplyr)
  library(tidyr)
  library(tibble)
  library(magrittr)

@@

  # Assign path based on computer name
  data_path <- file.path('c:', 'dropbox', 'research', 'airBNB', 'data')

  ## Load sources
  code_path <- file.path('c:', 'code', 'research', 'airBNBmelbourne')

  source(file.path(code_path, 'functions', "abb_Functions.R"))

@@

  # Property Information
  str_raw <- read.csv(file.path(data_path, 'raw', 'melbproperties.csv'), header=T)
  str_raw2 <- read.csv(file.path(data_path, 'raw', 'melbproperties_to_mar2017.csv'), 
                        header=T)

  # Daily rental information
  daily_raw  <- read.csv(file.path(data_path, 'raw', 'melbDaily.csv'), header=T)
  daily_raw2  <- read.csv(file.path(data_path, 'raw', 'melbDaily_to_mar2017.csv'), 
                           header=T)

@@

We then combine the two files for each type, standardize the field names and select only those fields which we need going forward.

@@

  # Property Level
  str_df <- dplyr::bind_rows(str_raw, str_raw2) %>%
    dplyr::select(property_id=Property.ID, host_id=Host.ID, listing_title=Listing.Title, 
                  property_type=Property.Type, listing_type=Listing.Type,
                  created_date=Created.Date, latitude=Latitude, longitude=Longitude,
                  rating=Overall.Rating, max_guests=Max.Guests, 
                  canc_policy=Cancellation.Policy, cleaning_fee=Cleaning.Fee, 
                  nightly_rate=Published.Nightly.Rate, 
                  weekly_rate=Published.Weekly.Rate, monthly_rate=Published.Monthly.Rate, 
                  avg_nightly_rate=Average.Daily.Rate, min_stay=Minimum.Stay, 
                  bedrooms=Bedrooms, bathrooms=Bathrooms)

  # Clean up
  rm(str_raw); rm(str_raw2)

  # Daily Data
  daily_df <- dplyr::bind_rows(daily_raw2, daily_raw) %>%
    dplyr::select(property_id=Property.ID, date=Date, status=Status, price=Price,
                  book_date=Booked.Date, resv_id=Reservation.ID)

  # Clean up
  rm(daily_raw); rm(daily_raw2);gc()

@@

Due to our combining of data, we have many duplicate observations.  We new remove any duplicates.  Note that the daily data in its raw format does not have a unique identifier, so we much first create one by combining the property id with the date of the observation.

@@

  # Structure Data
  str_df <- str_df %>% dplyr::filter(!duplicated(property_id))

  # Daily Data
  daily_df <- daily_df %>%
    dplyr::mutate(uniq_id = paste0(property_id, '_', date)) %>%
    dplyr::filter(!duplicated(uniq_id))

@@

 saveRDS(str_df, file=file.path(data_path, 'prepared', 'str_data.RDS'))

>>>>

---
title: "Pre Processing of the Short Term Data"
subtitle:  "For the 'To Airbnb' paper"
author: "Andy Krause and Gideon Aschwanden"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

This script executes a number of very long running data preparation process necessary to analyze the short term rental (Airbnb) data.  If you re-knit this file the code will not evaluate.  To actually evaluate the code you must manually re-run the code chunks individually in RStudio.

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

@@

```{r load_libraries, warnings=FALSE, message=FALSE}

  library(plyr)
  library(stringr)
  library(lubridate)
  library(dplyr)
  library(multidplyr) # Note you have to instal from the github site hadley/multidplyr
  library(tidyr)
  library(tibble)
  library(magrittr)

@@

  source(file.path(getwd(), 'functions', "abb_Functions.R"))

@@

  # Property information
  str_prop_1 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_prop_1.csv'), header=T)
  str_prop_2 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_prop_2.csv'), header=T)

  # Daily rental information
  str_daily_1 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_daily_1.csv'), header=T)
  str_daily_2 <- read.csv(file.path(getwd(), 'data', 'raw', 'str_daily_2.csv'), header=T)

@@

We then combine the two files for each type, standardize the field names and select only those fields which we need going forward.  We add a data origin field to the property level information so that we can take the most recent observation in later de-duplication processes.

@@

  # Property data
  str_df <- dplyr::bind_rows(str_prop_1 %>% 
                               dplyr::mutate(origin = 1), 
                             str_prop_2 %>% 
                               dplyr::mutate(origin = 2)) %>%
    dplyr::select(property_id=Property.ID, host_id=Host.ID, listing_title=Listing.Title, 
                  property_type=Property.Type, listing_type=Listing.Type,
                  created_date=Created.Date, latitude=Latitude, longitude=Longitude,
                  rating=Overall.Rating, max_guests=Max.Guests, 
                  canc_policy=Cancellation.Policy, cleaning_fee=Cleaning.Fee, 
                  nightly_rate=Published.Nightly.Rate, 
                  weekly_rate=Published.Weekly.Rate, monthly_rate=Published.Monthly.Rate, 
                  avg_nightly_rate=Average.Daily.Rate, min_stay=Minimum.Stay, 
                  bedrooms=Bedrooms, bathrooms=Bathrooms, origin)

  # Clean up
  rm(str_prop_1); rm(str_prop_2)

  # Daily data
  daily_df <- dplyr::bind_rows(str_daily_1, str_daily_2) %>%
    dplyr::select(property_id=Property.ID, date=Date, status=Status, price=Price,
                  book_date=Booked.Date, resv_id=Reservation.ID)

  # Clean up
  rm(str_daily_1); rm(str_daily_2);gc()

@@

Due to our combining of data, we have many duplicate observations.  We now remove any duplicates.  Note that the daily data in its raw format does not have a unique identifier, so we much first create one by combining the property id with the date of the observation.

@@

  # Property data
  str_df <- str_df %>% 
    dplyr::arrange(desc(origin)) %>%
    dplyr::filter(!duplicated(property_id)) %>%
    dplyr::select(-origin)

  # Daily data
  daily_df <- daily_df %>%
    dplyr::mutate(uniq_id = paste0(property_id, '_', date)) %>%
    dplyr::filter(!duplicated(uniq_id))

@@

 dir.create(file.path(getwd(), 'data', 'prepared'))
 saveRDS(str_df, file=file.path(getwd(), 'data', 'prepared', 'str_data.RDS'))
